LANGuage IDentification

### 2024-12-17 paper/paper.tex
basic understanding of back propagation

    \section*{License}
    \bigskip
    \begin{quote}
        \licensenotice
    \end{quote}
    \bigskip

@@

    \begin{equation}
        \begin{split}
            y_{k}(\mathbf{x}, \mathbf{w}) & = f_{n} \left( \displaystyle\sum_{i_{n}=0} w_{k i_{n}}^{(n)} \cdots f_{2} \left( \displaystyle\sum_{i_{2}=0} w_{i_{3}i_{2}}^{(2)} f_{1} \left( \displaystyle\sum_{i_{1}=0} w_{i_{2}i_{1}}^{(1)}x_{i_{1}} \right) \right) \cdots \right) = \\
            & = f_{n} \left( \cdots f_{2} \left( f_{1} \left( \mathbf{x} \cdot w^{(1)} \right)^{T} \cdot w^{(2)} \right)^{T} \cdots \cdot w^{(n)} \right) = \\
            & = Y_{n}
        \end{split}
        \label{eq:neural_network}
    \end{equation}

@@

>>>>

    \section{License}
    \bigskip
    \begin{quote}
        \licensenotice
    \end{quote}
    \bigskip

@@

    \begin{equation}
        \begin{split}
            y_{k}(\mathbf{x}, \mathbf{w}) & = f_{n} \left( \displaystyle\sum_{i_{n}=0} w_{k i_{n}}^{(n)} \hdots f_{2} \left( \displaystyle\sum_{i_{2}=0} w_{i_{3}i_{2}}^{(2)} f_{1} \left( \displaystyle\sum_{i_{1}=0} w_{i_{2}i_{1}}^{(1)}x_{i_{1}} \right) \right) \hdots \right) = \\
            & = f_{n} \left( \hdots f_{2} \left( f_{1} \left( \mathbf{x} \cdot w^{(1)} \right)^{T} \cdot w^{(2)} \right)^{T} \hdots \cdot w^{(n)} \right) = \\
            & = Y_{n}
        \end{split}
        \label{eq:neural_network}
    \end{equation}

@@

    \section{Learning}

    Now we can feed our network with some data, and it will give us a response
    (one that we might not want). The network will probably give you some
    nonsense output and that's probably because the weights in each neuron are
    also nonsense.

    Our goal now it's to try and find the optimal weights for the network.

    We will evaluate how bad the network is behaving and after that tell it
    what it should change to do a better job.

    \subsection{Error function}

    First we need a way of measuring how dumb the network is. Error functions
    measures the difference between the network output and the expected output.

    There are a lot of functions that we can choose to calculate that.

    We will use this $ E \left( \mathbf{w} \right) $ error function
    \eqref{eq:error_function} that is a sum of the squares of those differences
    in outputs.

    \begin{equation}
        E_{n} \left( \mathbf{w} \right) = \left| \mathbf{y}(\mathbf{x}_{n}, \mathbf{w}) - \mathbf{t}_{n} \right|^{2}
    \end{equation}

    \begin{equation}
        E \left( \mathbf{w} \right) = \frac{1}{2} \displaystyle\sum_{n=1}^{N} E_{n}(\mathbf{w})
        \label{eq:error_function}
    \end{equation}

    We have the input training matrix $ \left\{ \mathbf{x}_{n} \right\} $, $ N $
    is the input size. Also, we have $ \left\{ \mathbf{t}_{n} \right\} $ with the
    expected outputs for each input.

    Our goal is to minimize the error function \eqref{eq:error_function}
    because the closer to 0 that the error function return gets, means the
    differences between our neural network output and the expected output are
    less big.

    \subsection{Parameter Optimization}

    The way to minimize a function is by calculating its derivative on a point,
    so we can get the instantaneous rate of change at that point, that is, how
    much is the function growing/decreasing on that point. But that only works
    for unary function. Our neural network is a function with an unlimited
    (limited by the hardware which is running on) number of parameters (the
    weights). Because of that we instead use \textbf{partial derivatives}, and
    we calculate it for each one of the neural network weights (and biases if
    they are separated from the weights). Actually, we are calculation what's
    called the \textbf{gradient}, which is a vector of which one of those
    partial derivatives. That vector's direction is towards where that function
    grows (greatest rate of increase), we want to go the other way around.

    Our goal is to reach a minimum with $ \nabla E \left( \mathbf{w} \right) =
    \vec{0} $ and while we don't reach it, we keep changing the weights based
    on $ -\nabla E \left( \mathbf{w} \right) $. One of the problems that might
    arise is that $ \nabla E \left( \mathbf{w} \right) = \vec{0} $ does not
    mean that we found the perfect weight values, because a function can have
    multiple minimums and with the amount of parameters that a neural network
    has, it surely does, but also, it does not mean for sure we meet a minimum
    and instead meet a maximum of saddle point. Also targeting non error can be
    impossible, so we might instead target a maximum error that we allow the
    neural network to have.

    It also probably hard as hell to find the perfect weights only using math;
    I wouldn't get involved in that. At least it would take some time. So
    normally we just give each weight a random value, and we have the initial
    weight vector $ \mathbf{w}^{\left( 0 \right)} $. From that, we iteratively
    update the weights based on a weight vector update $ \Delta
    \mathbf{w}^{\left( \tau \right)} $ as shown in \eqref{eq:weight_updates}.

    \begin{equation}
        \mathbf{w}^{\left( \tau + 1 \right)} = \mathbf{w}^{\left( \tau \right)} + \Delta \mathbf{w}^{\left( \tau \right)}
        \label{eq:weight_updates}
    \end{equation}

    In our case we are going to do \textbf{gradient descent optimization} so
    from \eqref{eq:weight_updates} we get
    \eqref{eq:gradient_descent_optimization}.

    \begin{equation}
        \begin{split}
            \mathbf{w}^{\left( \tau + 1 \right)} & = \mathbf{w}^{\left( \tau \right)} + \Delta \mathbf{w}^{\left( \tau \right)} = \\
            & = \mathbf{w}^{\left( \tau \right)} + \left( \left( \mathbf{w}^{\left( \tau \right)} + \eta \nabla E \left( \mathbf{w}^{\left( \tau \right)} \right) \right) - \mathbf{w}^{\left( \tau \right)} \right) = \\
            & = \mathbf{w}^{\left( \tau \right)} + \eta \nabla E \left( \mathbf{w}^{\left( \tau \right)} \right)
        \end{split}
        \label{eq:gradient_descent_optimization}
    \end{equation}

    The $ \eta > 0 $ and it's called \textbf{learning rate}, so we can define
    how fast we want the optimization to be based on the gradient.

    While implementing this, I had some problems.

    Sometimes the $ E \left( \mathbf{w} \right) $ was minimizing and from out of nowhere it
    would explode into big numbers. But it seems there are some techniques that
    seem to maybe solve that problem such as \textbf{regularization} which is
    mentioned in page 30 \cite{bishop}. It can also be because of the way I
    calculated the derivatives: by using the limit definition. You can define
    an $ \epsilon $ on your neural network, and it will calculate the
    derivatives using it like this: $ \lim_{\epsilon\to\infty} \frac{f(x +
    \epsilon) - f(x)}{\epsilon} $. It would also depend on the input, and the
    initial weights.

    During implementation, I also thought that this system can cause a lot of
    problems in multilayer neural networks, because changing the weights of one
    neuron will affect the behavior of every neuron forward because they depend
    on its activation, and we are changing every weight at the same time.

    But I guess that's why \textbf{error backpropagation} is preferred over
    this method.

    \subsection{Error Backpropagation}

    Maybe because of what I mentioned above, error backpropagation exist, where
    the gradient is calculated from the last layer of the neural network (the
    output layer) to the first.

    This should work for any feed-forward neural network with any activation
    functions and error function.

    We have a general error function which is the sum of the errors of each
    input data.

    \begin{equation}
        E \left( \mathbf{w} \right) = \displaystyle\sum_{n=1}^{N} E_{n}(\mathbf{w})
    \end{equation}

    We want to compute $ \nabla E_{n} \left( \mathbf{w} \right) $.

    So we start by feeding forward where we calculate all the activations for
    every input $ \mathbf{x}_{i} $. In all the equations forward $ \left( \tau
    \right) $ represents the layer.

    Remember from the equation \eqref{eq:linear_model}. Here we have

    \newcommand{\nlayer}{\left( \tau \right)}
    \newcommand{\nplayer}[1]{\left( \tau - {#1} \right)}
    \newcommand{\npblayer}[1]{\left( \tau + {#1} \right)}
    \begin{equation}
        a_{k}^{\nlayer} = \displaystyle\sum_{i} \left( w_{ki}^{\nlayer} z_{i}^{\nplayer{1}} \right)
    \end{equation}

    \begin{equation}
        z_{i}^{\left( \tau \right)} = f^{\left( \tau \right)} \left( a_{i}^{\left( \tau \right)} \right)
    \end{equation}

    \newcommand{\pd}[2]{\frac{\partial {#1}}{\partial {#2}}}
    \newcommand{\ajl}{a_{j^{\nlayer}}^{\nlayer}}
    \newcommand{\ajpl}[1]{a_{j^{\npblayer{#1}}}^{\npblayer{#1}}}
    Now as we did before, we want to calculate the gradient, so we need to
    calculate the partial derivative $ \pd{E_{n}}{w_{ji}^{\nlayer}} $ for every
    weight.

    We can use the derivatives chain rule to write the partial derivative
    because the error function needs the activation after the weight to be
    dependent on that weight.

    \begin{equation}
        \pd{E_{n}}{w_{ji}^{\nlayer}} & = \pd{E_{n}}{a_{j}^{\nlayer}} \pd{a_{j}^{\nlayer}}{w_{ji}^{\nlayer}}
    \end{equation}

    And calculate those two separately.

    \begin{equation}
        \begin{split}
            \pd{E_{n}}{a_{j}^{\nlayer}} & = \pd{}{a_{j}^{\nlayer}} \left( a_{j}^{\left( \tau \right)} - t_{j} \right)^{2} = \\
            & = 2 \left( a_{j}^{\left( \tau \right)} - t_{j} \right) \pd{}{a_{j}^{\nlayer}} \left( a_{j}^{\left( \tau \right)} - t_{j} \right) = \\
            & = 2 \left( a_{j}^{\left( \tau \right)} - t_{j} \right) \left( 1 \right) = \\
            & = 2 \left( a_{j}^{\left( \tau \right)} - t_{j} \right) = \\
            & = \delta_{j}
        \end{split}
    \end{equation}
    \begin{equation}
        \begin{split}
            \pd{a_{j}^{\nlayer}}{w_{ji}^{\nlayer}} & = \pd{}{w_{ji}^{\nlayer}} \displaystyle\sum_{i} \left( w_{ji}^{\nlayer} z_{i}^{\nplayer{1}} \right) = \\
            & = \pd{}{w_{ji}^{\nlayer}} \left( w_{ji}^{\nlayer} z_{i}^{\nplayer{1}} + \displaystyle\sum_{k \neq i} \left( w_{jk}^{\nlayer} z_{k}^{\nplayer{1}} \right) \right) = \\
            & = \left( 1 \right) z_{i}^{\nplayer{1}} + \displaystyle\sum_{k \neq i} 0 = \\
            & = z_{i}^{\nplayer{1}}
        \end{split}
    \end{equation}

    \begin{equation}
        \pd{E_{n}}{w_{ji}} = \delta_{j}^{\nlayer} z_{i}^{\nplayer{1}}
    \end{equation}

    We call $ \delta_{j}^{\nlayer} $ the errors. Because we already calculated
    every $ z_{i}^{\nplayer{1}} $ while forward propagating, we really just
    need to calculate the errors.

    The $ \delta_{j}^{\nlayer} $ we calculated is for the output layer. For the
    layer before that, we can also use the chain rule.

    \begin{equation}
        \begin{split}
            \delta_{j}^{\nplayer{1}} & = \pd{E_{n}}{a_{j}^{\nplayer{1}}} \\
            & = \displaystyle\sum_{k} \left( \pd{E_{n}}{a_{k}^{\nlayer}} \pd{a_{k}^{\nlayer}}{a_{j}^{\nplayer{1}}} \right) = \\
            & = \displaystyle\sum_{k} \left( \delta_{k}^{\nlayer} \pd{}{a_{j}^{\nplayer{1}}} \left( w_{kj}^{\nlayer} f^{\nplayer{1}} \left( a_{j}^{\nplayer{1}} \right) + \displaystyle\sum_{i \neq j} \left( w_{ki}^{\nlayer} z_{ki}^{\nplayer{1}} \right) \right) \right) = \\
            & = \displaystyle\sum_{k} \left( \delta_{k}^{\nlayer} w_{ki}^{\nlayer} f^{\prime \nplayer{1}} \left( a_{i}^{\nplayer{1}} \right)  \right) = \\
            & = f^{\prime \nplayer{1}} \left( a_{j}^{\nplayer{1}} \right) \displaystyle\sum_{k} \left( w_{kj}^{\nlayer} \delta_{k}^{\nlayer} \right)
        \end{split}
    \end{equation}

    \begin{equation}
        \begin{split}
            \pd{E_{n}}{w_{ji}^{\nlayer}} & = \pd{E_{n}}{a_{j}^{\nlayer}} \pd{a_{j}^{\nlayer}}{w_{ji}^{\nlayer}} = \\
            & = \pd{E_{n}}{\ajpl{l}} \displaystyle\prod_{L = 1}^{l} \left( \pd{\ajpl{l - \left( L - 1 \right)}}{\ajpl{l - L}} \right) \pd{\ajl}{w_{ji}^{\nplayer{l}}}
        \end{split}
    \end{equation}

    After calculating all errors from the output layer, you are able to
    calculate from the other layers, one at a time, backwards, and you'll end
    up with the derivatives for the gradient.

    We may also accumulate gradients to apply it to the weights all at once.

    \begin{equation}
        \pd{E}{w_{ji}^{\nlayer}} & = \displaystyle\sum_{n} \pd{E_{n}}{w_{ji}^{\nlayer}}
    \end{equation}
